---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Autostats
  
## Overview

The goal of autostats is to help you perform basics __medical statistics__ from a dataframe, such as :

1. **Table 1** : percentage, median, mean and statistical test between groups (khi2, Fisher, student)
1. **Logistic regression** : with automated multivariate variable selection

## Installation

```{r, eval = FALSE}
# The package is only available from github :
# install.packages("devtools")
devtools::install_github("TanguyPerennec/Autostats")
```


## Getting started

```{r}
library(Autostats)
```

Autostats functions are :

* **"Table 1"** which render an usual table 1 with basics descriptive statistics (median, mean and proportions) of 2 groups (define by a binary variable called "y") from a dataframe (called "DF"). Use `Table1(DF,y)`
  
* **"Logistic regression"** which render a matrix with the results (odds ratios, confidence intervals and p-values) of univariate logistic regression of all the variables provided from the dataframe, and the results of a multivariate logistic regression with automated variable selection. Use `reglog(DF,y)` to have the matrix results or just `multivariate_selection(DF,y)` to have the selection with details of the selection method. You can provide a different method ("backward","forward","stepwise","augmented") and selection criteria ("deviance", "AIC", "BIC").


## Related work

Autostats will be used in an simulation project of multivariate logistic regression selection so as to provide a best understanding of multivariate selection process. 

---
